---
title: "GHRU test"
author: "Ojonugwa Abubakar"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: cayman
    fig_width: 15
    fig_height: 15
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, tidy = TRUE, comment = "", dev = "svg")
```

# Loading the packages
```{r load packages}
library(prettydoc)
library(knitr)
library(ggplot2)
library(tidyverse)
library(scales)
library(formatR)
library(printr)
```

# Importing the data
```{r import data}
data <- read.csv("/Users/otabu/OneDrive/Desktop/GHRU data.csv", header = T, check.names = F)
```

# Question 1
```{r Q1}
## Samples that have failed the contamination check
data %>%
  filter(confindr.contam_status.check_result == "FAILURE") %>%
  count()

## Samples that failed the contamination check and have contamination"
data %>%
  filter(confindr.contam_status.check_result == "FAILURE" & 
      confindr.percentage_contamination.metric_value > 5.0) %>%
  count()
```
57 samples have contamination and failed the contamination check.

# Question 2
```{r Q2}
## obtaining the names of the columns in the data
colnames(data)

## samples with less than or equal to 50 contigs and a N50 value of greater than or equal to 750,000
data %>%
  filter(`quast.# contigs (>= 1000 bp).metric_value` <= 50 & 
      quast.N50.metric_value >= 750000) %>%
  count()
```
49 samples have less than or equal to 50 contigs and a N50 value of greater than or equal to 750,000

# Question 3
```{r Q3}
## selecting only columns with numeric columns and renaming them
data_num <- data %>%
  select(1, where(is.numeric)) %>%
  rename_with(stringr::str_replace,
              pattern = "quast.", replacement = "", matches ("quast.")) %>%
  rename_with(stringr::str_replace,
              pattern = ".metric_value", replacement = "", matches (".metric_value")) %>%
  rename(contamination_percent = confindr.percentage_contamination) 

## obtaining names of column of data
colnames(data_num)
```

# Question 4
```{r Q4}
## range of total length
range(data_num$`Total length (>= 1000 bp)`)

## Boxplot of Total length (>= 1000bp)
ggplot(data_num) +
  geom_boxplot(aes(`Total length (>= 1000 bp)`)) +
  scale_x_continuous(breaks = pretty_breaks(n = 16), 
                     minor_breaks = seq(6000000, 14000000, by = 500000), labels = label_number(suffix = "", scale = 1e-6)) +
  labs(title = "Box blot showing the distribution of total length (>= 1000 bp) from a QC spreadsheet",
       x = "Total length (in millions)") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        axis.title = element_text(size=12), 
        plot.title = element_text(size=15))
```
The box plot shows that while the median total length (>=1000 bp) is around 7,000,000 bp, there are outlier values of up to 13,500,000. Hence, the boxplot is skewed to the right.

# Question 5
```{r Q5}
## tidying the data
data_pivot <- data_num %>%
  pivot_longer(!sample_name, names_to = "metric", values_to = "value")

## printing the first 20 rows of the tidy table
head(data_pivot, 20)
```

# Question 6
```{r Q6}
## Violin plot of numeric variables with jittered data points
ggplot(data_pivot, aes(metric, value)) +
  geom_violin() +
  geom_jitter(position = position_jitter(seed = 1, width = 0.1)) +
  scale_y_continuous(trans = 'log2') +
  theme(axis.text.x = element_text(angle = 60, hjust = 1, size=10), 
        axis.text.y = element_text(size=10), 
        axis.title = element_text(size=12), 
        plot.title = element_text(size=15)) + 
  ggtitle("Violin plot showing distribution of numerical variables in a QC spreadsheet")
```

* The violin plot shows:
  + a lot of strains have no contamination, with contaminated strains having less than 60% contamination. 
  + N50 distribution is left skewed, with few strains having smaller N50 values. 
  + The values of the total length are clustered close to the median. The distribution is not skewed as there are no distinct outliers.


